### 2022-12-06 StatsProject_draft_Caro.Rmd
Update Carolina branch with StatsProject_draft_Caro.Rmd

---
title: "Statistics Project"
output: html_notebook
---

# ANALYSIS 1

Goal: Estimate how SalesPrice of the house is related to the square footage of the living area of the house & if SalesPrice (and its relationship to square footage) depends on which neighborhood the house is located in. We are only interested at looking at houses in the NAmes, Edwards & BrkSide neighborhoods.

## Packages
```{r}
library(dplyr)
library(readr)
library(tidyverse)
```

## Read in data
```{r}
train <- read.csv("train.csv")
test <- read.csv("test.csv")

head(train)
head(test)
```
## Filter data on neighborhoods 
```{r}
train_a1 <- dplyr::filter(train,Neighborhood =="Edwards" | Neighborhood =="NAmes" | Neighborhood == "BrkSide")
test_a1 <- dplyr::filter(test,Neighborhood =="Edwards" | Neighborhood =="NAmes" | Neighborhood == "BrkSide")

head(train_a1)
head(test_a1)
```
## Summary Statistics
```{r}
summary(train_a1)
summary(test_a1)
```
## EDA 
```{r}
train_a1 %>% ggplot(aes(x = (GrLivArea), y = SalePrice/1000, color = Neighborhood)) + geom_point() + stat_smooth(method = "lm")

train_a1 %>% ggplot(aes(x = SalePrice/10000, fill = Neighborhood)) + geom_histogram(bins =20)

train_a1 %>% ggplot(aes(x = GrLivArea, fill = Neighborhood)) + geom_histogram(bins =20)
```
## Log Transforms 
```{r}
# histograms with log transform on each individual variables
train_a1 %>% ggplot(aes(x = log(SalePrice/10000), fill = Neighborhood)) + geom_histogram(bins =20)

train_a1 %>% ggplot(aes(x = log(GrLivArea), fill = Neighborhood)) + geom_histogram(bins =20)

train_a1 %>% ggplot(aes(x = log(GrLivArea), y = SalePrice/1000, color = Neighborhood)) + geom_point() + stat_smooth(method = "lm")

train_a1 %>% ggplot(aes(x = (GrLivArea), y = log(SalePrice/1000), color = Neighborhood)) + geom_point() + stat_smooth(method = "lm")

train_a1 %>% ggplot(aes(x = log(GrLivArea), y = log(SalePrice/1000), color = Neighborhood)) + geom_point() + stat_smooth(method = "lm")
```

# Linear Model
```{r}
fit <- lm(log(SalePrice) ~ log(GrLivArea) + Neighborhood + GrLivArea*Neighborhood, data = train_a1)
summary(fit)
confint(fit)

plot(fit)
```
# ANOVA 
```{r}
res.aov <- aov(log(SalePrice)~log(GrLivArea)+Neighborhood+Neighborhood*GrLivArea, data = train_a1)
summary(res.aov)
```
```{r}
library(car)
# QQPlot
qqPlot(log(train_a1$SalePrice))

# histogram of residuals 
h <- hist(fit$residuals, breaks = 10, density = 10,col = "lightgray",main = "Residuals") 
xfit <- seq(min(fit$residuals), max(fit$residuals), length = 40) 
yfit <- dnorm(xfit, mean = mean(fit$residuals), sd = sd(fit$residuals)) 
yfit <- yfit * diff(h$mids[1:2]) * length(fit$residuals) 
```